Analysis of Open Data and Computational Reproducibility in Registered Reports in Psychology

### 2020-01-23 manuscript_version_2/reproducing_registered_reports.tex
revision after co-author feedback

\abstract{
Ongoing technological developments have made it easier than ever before for scientists to share their data, materials, and analysis code. Sharing data and analysis code makes it easier for other researchers to re-use or check published research. These benefits will only emerge if researchers can reproduce the analysis reported in published articles, and if data is annotated well enough so that it is clear what all variables mean. Because most researchers have not been trained in computational reproducibility, it is important to evaluate current practices to identify practices that can be improved. We examined data and code sharing, as well as computational reproducibility of the main results, without contacting the original authors, for Registered Reports published in the psychological literature between 2014 and 2018. Of the main results from 62 articles that met our inclusion criteria, data was available for 41 articles, and analysis scripts for 37 articles. For the main results in 36 articles that shared both data and code and performed analyses in SPSS, R, Python, MATLAB, or JASP, we could run the scripts for 31 analyses, and reproduce the main results for 20 articles. Although the articles that shared both data and code (36 out of 62, or 58\%) and articles for which main results could be computationally reproduced (21 out of 36, or 58\%) was relatively high compared to other studies, there is clear room for improvement. We provide practical recommendations based on our observations, and link to examples of good research practices in the papers we reproduced.

@@

Researchers are currently exploring ways to make science more open and transparent. Among novel developments such as pre-registration, preprints, and open peer review, an increasing number of journals, funders, and researchers are beginning to expect that data, materials, and analysis code will be shared by default with scientific publications (e.g., (Morey et al., 2016). Sharing data and analysis code with scientific publications allows others to more easily reproduce, check, and build on existing work. This requires the development of new skills and best practices, since most scientists have not received training in how to make their work reproducible. It is important to evaluate how data and code are currently being shared, and how easy it is to reproduce analyses reported in the published literature, to learn what can be improved. With this goal in mind, we computationally reproduced the main results of Registered Reports published in the psychological literature.

@@

Recently, scholars have started to empirically examine the extent to which data is shared with published articles, and, if so, whether it was possible to reproduce the data analysis reported in the articles. Hardwicke et al. (2018) examined the analytic reproducibility of 35 articles published in the journal \emph{Cognition}. Target outcomes that supported the identified substantive finding of eleven articles could be reproduced by independently writing analysis code, without assistance from the original authors, and 13 sets of target outcomes contained at least one outcome that could not be reproduced even with author assistance. The authors estimated that it took between 2-25 hours to reproduce the reported results per article, but they did not record the exact time. Stockemer, Koehler, and Lentz (2018) analyzed reproducibility in all articles published in 2015 in three political science journals. They emailed authors for the code and data, and found that for 71 articles for which they received code and data, one paper's results could not be reproduced due to a lack of a software license, and 16 articles' findings could not be reproduced with access to the required software. For the remaining articles, 32 sets of results could be exactly reproduced, 19 could be reproduced with slight differences, and 3 articles yielded significantly different results. Stodden, Seiler, and Ma (2018) analyzed data availability in the journal \emph{Science} in 2011-2012 and found that 26 of 204 (or 13\%) of articles provided information to retrieve data and/or code without contacting the authors. For all datasets they acquired after e-mailing authors for data and code, 26\% were estimated to be computationally reproducible. These studies reveal that there is clear room for improvement in how reproducible published articles are.

We set out to examine the data availability and reproducibility in Registered Reports published in psychological science. Our main interest was to examine the computational reproducibility of the main analyses reported in published articles, without contacting the original authors. One of the main benefits of sharing data and code alongside an article (compared to making these files available upon request) is that results can be reproduced, and data reused even if the original author can no longer be reached.

@@

The novelty of Registered Reports may attract early adopters who are also exploring other novel developments aimed at improving research practices in psychology, such as data and code sharing. In addition, one journal that publishes Registered Reports (Royal Society Open Science) requires authors to deposit data and code, and several other journals that publish Registered Reports \enquote{strongly encourage} data and/or code sharing. We therefore hypothesized that researchers who publish Registered Reports would also be likely to share data and code in public repositories, and to embrace computational reproducibility. To evaluate the reproducibility of findings published in Registered Reports, we examined if data could be located, were available at the indicated location, could be opened in open-source or accessible software, were documented well enough to be understandable, and could be used to reproduce the main analyses reported in the published manuscript.

@@

To find Registered Reports published in psychology, we drew from a database of registered reports maintained by the Center for Open Science.\footnote{\url{https://www.zotero.org/groups/479248/osf/items/collectionKey/KEJP68G9}} At the start of this project (06-05-2018) this database consisted of 118 published Registered Reports. Seventy-nine articles in this database were published in the psychological literature. We limited our analysis to studies performed by single groups, since such articles are most representative of the researchers currently work, and excluded 7 large scale collaborations where dedicated team members were responsible for making the analysis reproducible, such as Registered Replication Reports (Hagger et al., 2016) and Many Labs projects (Klein et al., 2014; R. A. Klein et al., 2018). Upon further inspection of the 72 papers left in the dataset, it turned out that 10 were not formally Registered Reports. This left 62 studies in our sample. When evaluating whether we could reproduce the original results, we limited ourselves to statistical software packages that we had experience with (R, SPSS, Python, MATLAB, and JASP) and excluded studies that required expertise in software packages we are not trained in (e.g., dedicated EEG software), which led to the exclusion of one additional study.

@@

We set out to reproduce the findings of 62 papers that met our inclusion criteria. For each article, we coded whether the data and code were (a) linked, (b) available, (c) not software-specific, (d) understandable, and (e) reproducible. These five categories were inspired by FAIR data principles (Wilkinson et al., 2016) that address the findability, accessibility, interoperability, and reusability of data and code, but we did not explicitly code whether papers adhered to the exact definitions of the four FAIR principles. Our main aim was to examine the reproducibility of results, while adhering to FAIR criteria requires meeting more stringent requirements, for example concerning the presence of metadata (which was missing for practically all datasets). We considered data and code to be linked when the manuscript included a unique link to the data and scripts. Ideally, such a link consists of a stable digital object identifier (DOI). A hyperlink to a website that contains the data and script also suffices, although hyperlinks are known to break over time (Gertler \& Bullock, 2017).

Of the 62 papers, 45 linked to data and/or code that were part of the article. Linking to data or code does not mean the data is actually available, and not linking to data in an article does not mean data is unavailable. For one article the link no longer worked (highlighting the benefit of using a stable DOI for linking to data). For three articles, the link still worked, but there was no data at the linked destination. For three articles where there was no link to the data in the article, we were nevertheless able to find data on the Open Science Framework when searching for the title of the paper. For one paper the data was linked but not available, as the data was embargoed until a future date.

@@

When coding whether the script could be executed and the results could be reproduced we resorted to a dichotomous classification (\enquote{yes} or \enquote{no}), but coders often reported \enquote{partial} reproducibility. Code often needed minor adjustments to run on the data, such as changing file locations, or loading packages in R, and coders sometimes took different approaches to how much they would adjust the code to make it run on the data (for detailed comments, see our data file). When judging if the code ran on the data, we allowed for minor errors, but categorized code as not running when it was unclear how analysis code related to data files, or when there were a substantial number of errors when attempting to run all the code, even after attempts to make minor corrections. Furthermore, we did not preregister this study, and had no clearly defined coding scheme based on pilot data, and as a consequence coders initially used different thresholds of reproducibility, based on whether every single result reported in a paper could be reproduced by the code and data, or only the main results reported in the article. After evaluating our initial coding round, we considered an article reproducible when we could get the same \emph{main results} as represented in the paper with at best \emph{minor changes} to the analysis scripts. This means we considered the analysis reproducible, even if in the absence of a codebook coders had to search through the analysis code and dataset to identify how variable and label names related to the results reported in the paper. Furthermore, we changed folder locations when needed, and installed and/or loaded required libraries. Finally, even though sometimes figures were generated in R and contained relevant information (e.g., the pattern of means) we did not require all figures to be reproducible. For each study reported in the article we identified main results (i.e., any reported descriptives and statistical tests) based on the research question highlighted in the title and abstract. Approximately half of the Registered Reports consisted of replication studies, where the main analysis was explicitly stated based on a previous study. In the remaining Registered Reports the main research question was often stated clearly in a \enquote{confirmatory analyses} section. Nevertheless, some arbitrary judgments were required when deciding which analyses were part of the main results.

There were differences between coders in how much expertise they had with R, SPSS, and JASP (PO had less experience, while JG, NC and DL had more experience; SG ran all MATLAB and Python code). The more expertise coders had, the easier it was to reproduce findings (which lead to lower inter-rater reliability). This raises the question regarding which level of skill is expected to be able to reproduce results reported in a scientific article. The required experience might be difficult to quantify. Our results concerning the reproducibility of results is based on which results a PhD student with experience in the statistical software and educated in the same scientific discipline could reproduce, which we feel is a reasonable standard. For the current analysis, we considered a study reproducible as long as the more experienced coder could reproduce the main results. All authors have collectively discussed disagreements, which on several occasions led to clarifying ambiguities in rating strategies and correcting mistakes in the process of paper analysis (e.g.~overlooked script files). The final ratings presented here have been discussed and approved by all authors, but will not be completely flawless, and we expect different teams of coders to reach slightly different conclusions (i.e., perfect reliability would be extremely difficult to achieve). It may bear repeating that the main goal of our analysis is to evaluate where there is room for improvement, and identify practices that researchers can use to make their work more reproducible.

@@

For the 15 articles that could not be reproduced, the main reasons were that 1) code to reproduce some values was missing (e.g., dedicated code to run the a macro in SPSS), which occurred 8 times, 2) code gave errors (e.g., because variables in the dataset were missing, or because functions did not run as expected), which occurred 6 times, and 3) in one case the code might have been reproducible, but the code was so complex that after 40 minutes only a small part of the results could be reproduced, and it was judged that the time needed to reproduce the results fell outside of a reasonable time span. Two small errors were observed (one value of 0.89 was rounded to 0.88, and probable type where a value of 0.06 should have been 0.03) but since dozens of other values in these articles reproduced perfectly, these errors were not considered severe enough to deem the main results not computationally reproducible.

After the data and code had been downloaded, and the reported results were identified by looking through the article, the average time to reproduce analyses reported in R was 27.08 minutes for the first coder (SD = 28.55) and 32.50 minutes (SD = 20.95) for the second coder. Reproducing the SPSS analysis took on average 17.35 minutes (SD = 9.54) for the first coder and 25.50 minutes (SD = 9.72) for the second coder. Most of the time was spent on matching output from the statistical analyses to the analyses reported in the manuscript. This suggests that even if results are reproducible, the organization of the output, and the relation of the output to the published manuscript, can often be improved.

@@

Nevertheless, the reproducibility rate for articles where data and code were available show there is clear room for improvement in how reproducible published Registered Reports are, when we attempted to reproduce main analyses without contacting the authors. One of the main goals of our project was to identify ways to improve the reproducibility of published articles. We encountered several common issues that made results reported in Registered Reports difficult to reproduce (cf.~Hardwicke et al., 2018). This leads to 4 points researchers in psychology should focus on to improve reproducibility, namely (a) add a codebook to data files, (b) annotate code so it is clear what the code does, and clearly structure code (e.g., using a README) so others know which output analysis code creates, (c) check whether the code you shared still reproduces all analyses after revisions during the peer review process, and (d) list packages that are required in R and the versions used at the top of your R file. We will discuss each of these points below, and link to examples of good practices that we encountered.

@@

In addition to novel technologies, most progress can probably be made by developing standards within research communities, and educating researchers about best practices that guarantee reproducibility (for recent examples, see {[}O. Klein et al. (2018); liu\_successes\_2019{]}. Most researchers are not trained in reproducible data analysis, and cannot be expected to invent best practices from scratch. As good examples appear in the published literature over time, and best practices within subdisciplines crystalize, standards should emerge that improve reproducibility, and that allow researchers to share data and code in such a way that others with basic scientific training can reproduce their results and reuse their data.

@@

>>>>

\abstract{
Ongoing technological developments have made it easier than ever before for scientists to share their data, materials, and analysis code. Sharing data and analysis code makes it easier for other researchers to re-use or check published research. However, these benefits will only emerge if researchers can reproduce the analysis reported in published articles and if data is annotated well enough so that it is clear what all variables mean. Because most researchers are not trained in computational reproducibility, it is important to evaluate current practices to identify practices that can be improved. We examined data and code sharing for Registered Reports published in the psychological literature between 2014 and 2018, and attempted to independently computationally reproduce the main results in each article. Of the main results from 62 articles that met our inclusion criteria, data were available for 41 articles, and analysis scripts for 37 articles. For the main results in 36 articles that shared both data and code we could run the scripts for 31 analyses, and reproduce the main results for 20 articles. Although the articles that shared both data and code (36 out of 62, or 58\%) and articles for which main results could be computationally reproduced (21 out of 36, or 58\%) was relatively high compared to other studies, there is clear room for improvement. We provide practical recommendations based on our observations and link to examples of good research practices in the papers we reproduced.

@@

Researchers are currently exploring ways to make science more open and transparent. Among novel developments such as pre-registration, preprints, and open peer review, an increasing number of journals, funders, and researchers are beginning to expect that data, materials, and analysis code will be shared by default with scientific publications (e.g., Morey et al. (2016)). Sharing data and analysis code with scientific publications allows others to more easily reproduce, check, and build on existing work. This requires the development of new skills and best practices, since most scientists have not received training in how to make their work reproducible. It is important to evaluate how data and code are currently being shared, and how easy it is to reproduce analyses reported in the published literature, to learn what can be improved. With this goal in mind, we computationally reproduced the main results of Registered Reports published in the psychology literature.

@@

Recently, scholars have started to empirically examine the extent to which data is shared with published articles, and, if so, whether it was possible to reproduce the data analyses reported in the articles. Hardwicke et al. (2018) examined the analytic reproducibility of 35 articles published in the journal \emph{Cognition}. Target outcomes that supported the identified substantive finding of eleven articles could be reproduced by independently writing analysis code, without assistance from the original authors, and 13 sets of target outcomes contained at least one outcome that could not be reproduced even with author assistance. The authors estimated that it took between 2-25 hours to reproduce the reported results per article, but they did not record the exact time. Stockemer, Koehler, and Lentz (2018) analyzed reproducibility in all articles published in 2015 in three political science journals. They emailed authors for the code and data, which they received for 71 articles. The results of one article could not be reproduced due to a lack of a software license, and 16 articles' findings could not be reproduced with access to the required software. For the remaining articles, 32 sets of results could be exactly reproduced, 19 could be reproduced with slight differences, and 3 articles yielded significantly different results. Stodden, Seiler, and Ma (2018) analyzed data availability in the journal \emph{Science} in 2011-2012 and found that 26 of 204 (or 13\%) of articles provided information to retrieve data and/or code without contacting the authors. For all datasets they acquired after e-mailing authors for data and code, 26\% were estimated to be computationally reproducible. These studies reveal that there is clear room for improvement in how reproducible published articles are.

We set out to examine the data availability and reproducibility in Registered Reports published in psychological science. Our main interest was to examine the computational reproducibility of the main analyses reported in published articles, without contacting the original authors. One of the main benefits of sharing data and code alongside an article (compared to making these files available upon request) is that results can be reproduced and data reused even if the original author can no longer be reached.

@@

The novelty of Registered Reports may attract early adopters who are also exploring other novel developments aimed at improving research practices in psychology, such as data and code sharing. In addition, one journal that publishes Registered Reports (Royal Society Open Science) requires authors to deposit data and code, and several other journals that publish Registered Reports strongly encourage data and/or code sharing. We expected researchers who publish Registered Reports to be likely to share data and code in public repositories, and to embrace computational reproducibility. To evaluate the reproducibility of findings published in Registered Reports, we examined if data could be located, were available at the indicated location, could be opened in open-source or accessible software, were documented well enough to be understandable, and could be used to reproduce the main analyses reported in the published manuscript.

@@

To find Registered Reports published in psychology, we drew from a database of registered reports maintained by the Center for Open Science.\footnote{\url{https://www.zotero.org/groups/479248/osf/items/collectionKey/KEJP68G9}} At the start of this project (06-05-2018) this database consisted of 118 published Registered Reports. Seventy-nine articles in this database were published in the psychology literature. We limited our analysis to studies performed by single groups, since such articles are most representative of the researchers currently work, and excluded 7 large scale collaborations where dedicated team members were responsible for making the analysis reproducible, such as Registered Replication Reports (Hagger et al., 2016) and Many Labs projects (Klein et al., 2014; R. A. Klein et al., 2018). Upon further inspection of the 72 papers left in the dataset, it turned out that 10 were not formally Registered Reports. This left 62 studies in our sample. When evaluating whether we could reproduce the original results, we limited ourselves to statistical software packages that we had experience with (R, SPSS, Python, MATLAB, and JASP) and excluded studies that required expertise in software packages we were not trained in (e.g., dedicated EEG software), which led to the exclusion of one additional study.

@@

We set out to reproduce the findings of 62 papers that met our inclusion criteria. For each article, we coded whether the data and code were (a) linked, (b) available, (c) not software-specific, (d) understandable, and (e) reproducible. These five categories were inspired by FAIR data principles that address the findability, accessibility, interoperability, and reusability of data and code, but we did not explicitly code whether papers adhered to the exact definitions of the four FAIR principles (Wilkinson et al., 2016). Our main aim was to examine the reproducibility of results, while adhering to FAIR criteria requires meeting more stringent requirements, for example concerning the presence of metadata (which was missing for all datasets). We considered data and code to be linked when the manuscript included a unique link to the data and scripts. Ideally, such a link consists of a stable digital object identifier (DOI). A hyperlink to a website that contains the data and script also suffices, although hyperlinks are known to break over time (Gertler \& Bullock, 2017).

Of the 62 papers, 45 linked to data and/or code that were part of the article. Linking to data or code does not mean the data is actually available, and not linking to data in an article does not mean data is unavailable. For one article the link no longer worked (highlighting the benefit of using a stable DOI for linking to data). For three articles, the link still worked, but there was no data at the linked destination. For three articles where there was no link to the data in the article, but we were nevertheless able to find data on the Open Science Framework when searching for the title of the paper. For one paper the data was linked but not available, as the data was embargoed until a future date.

@@

When coding whether the script could be executed and the results could be reproduced we used a dichotomous classification (\enquote{yes} or \enquote{no}), but coders often reported \enquote{partial} reproducibility. Code often needed minor adjustments to run on the data, such as changing file locations, or loading packages in R, and coders sometimes took different approaches to how much they would adjust the code to make it run on the data (for detailed comments, see our data file). When judging if the code ran on the data, we allowed for minor errors, but categorized code as not running when it was unclear how analysis code related to data files, or when there were a substantial number of errors when attempting to run all the code, even after attempts to make minor corrections. Furthermore, we did not preregister this study, and had no clearly defined coding scheme based on pilot data. As a consequence coders initially used different thresholds of reproducibility, based on whether every single result reported in a paper could be reproduced by the code and data, or only the main results reported in the article. After evaluating our initial coding round, we considered an article reproducible when we could get the same \emph{main results} as represented in the paper with at best \emph{minor changes} to the analysis scripts. This means we considered the analysis reproducible, even if in the absence of a codebook coders had to search through the analysis code and dataset to identify how variable and label names related to the results reported in the paper. Furthermore, we changed folder locations when needed, and installed and/or loaded required libraries. Finally, even though some figures were generated in R and contained relevant information (e.g., the pattern of means) we did not require all figures to be reproducible. For each study reported in the article we identified main results (i.e., any reported descriptives and statistical tests) based on the research question highlighted in the title and abstract. Approximately half of the Registered Reports consisted of replication studies, where the main analysis was explicitly stated based on a previous study. In the remaining Registered Reports the main research question was often stated clearly in a \enquote{confirmatory analyses} section. Nevertheless, some arbitrary judgments were required when deciding which analyses were part of the main results.

There were differences between coders in how much expertise they had with R, SPSS, and JASP (PO had less experience, while JG, NC and DL had more experience; SG ran all MATLAB and Python code). The more expertise coders had, the easier it was to reproduce findings (which lead to lower inter-rater reliability). This raises the question regarding which level of skill is expected to be able to reproduce results reported in a scientific article. The required experience might be difficult to quantify. Our results concerning the reproducibility of results is based on which results a PhD student with experience in the statistical software and educated in the same scientific discipline could reproduce, which we feel is a reasonable standard. For the current analysis, we considered a study reproducible as long as the more experienced coder could reproduce the main results. All authors collectively discussed disagreements, which on several occasions led to clarifying ambiguities in rating strategies and correcting mistakes in the process of paper analysis (e.g.~overlooked script files). The final ratings presented here have been discussed and approved by all authors, but will not be completely flawless, and we expect different teams of coders to reach slightly different conclusions (i.e., perfect reliability would be extremely difficult to achieve). It may bear repeating that the main goal of our analysis is to evaluate where there is room for improvement, and identify practices that researchers can use to make their work more reproducible.

@@

For the 15 articles that could not be reproduced, the main reasons were that 1) code to reproduce some values was missing (e.g., dedicated code to run a macro in SPSS), which occurred 8 times, 2) code gave errors (e.g., because variables in the dataset were missing, or because functions did not run as expected), which occurred 6 times, and 3) in one case the code might have been reproducible, but the code was so complex that after 40 minutes only a small part of the results could be reproduced, and it was judged that the time needed to reproduce the results fell outside of a reasonable time span. In the articles that were coded as reproducible, two small errors were observed (one value of 0.89 was rounded to 0.88, and probable type where a value of 0.06 should have been 0.03) but since dozens of other values in these articles reproduced perfectly, these errors were not considered severe enough to deem the main results not computationally reproducible.

After the data and code had been downloaded, and the reported results were identified by looking through the article, the average time to reproduce analyses reported in R was 27.08 minutes (SD = 28.55) for the first coder and 32.50 minutes (SD = 20.95) for the second coder. Reproducing the SPSS analysis took on average 17.35 minutes (SD = 9.54) for the first coder and 25.50 minutes (SD = 9.72) for the second coder. Most of the time was spent on matching output from the statistical analyses to the analyses reported in the manuscript. This suggests that even if results are reproducible, the organization of the output, and the relation of the output to the published manuscript, can often be improved.

@@

Nevertheless, our results indicate that that there is clear room for improvements in the computational reproducibility of Registered Reports. One of the main goals of our project was to identify ways to improve the reproducibility of published articles. We encountered several common issues that made results reported in Registered Reports difficult to reproduce (cf.~Hardwicke et al., 2018). This leads to 4 points researchers in psychology should focus on to improve reproducibility, namely (a) add a codebook to data files, (b) annotate code so it is clear what the code does, and clearly structure code (e.g., using a README) so others know which output analysis code creates, (c) check whether the code you shared still reproduces all analyses after revisions during the peer review process, and (d) list packages that are required in R and the versions used at the top of your R file. We will discuss each of these points below, and link to examples of good practices that we encountered.

@@

In addition to novel technologies, most progress can probably be made by developing standards within research communities, and educating researchers about best practices that guarantee reproducibility (for recent examples, see (O. Klein et al., 2018; Liu \& Salganik, 2019). Most researchers are not trained in reproducible data analysis, and cannot be expected to invent best practices from scratch. As good examples appear in the published literature over time, and best practices within subdisciplines crystalize, standards should emerge that improve reproducibility, and that allow researchers to share data and code in such a way that others with basic scientific training can reproduce their results and reuse their data.

@@

\leavevmode\hypertarget{ref-liu_successes_2019}{}%
Liu, D., \& Salganik, M. (2019). Successes and struggles with computational reproducibility: Lessons from the Fragile Families Challenge. doi:\href{https://doi.org/10.31235/osf.io/g3pdb}{10.31235/osf.io/g3pdb}
